---
title: "Using the sprintr package"
author: "Guo Yu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
#output: pdf_document
vignette: >
  %\VignetteIndexEntry{Using the sprintr package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `sprintr` package contains the implementations of a computationally efficient method, called sprinter, to fit large interaction models based on the reluctant interaction selection principle. The details of the method can be found in [Yu, Bien, and Tibshirani (2019) *Reluctant interaction modeling*](https://arxiv.org/abs/1907.08414).
In particular, `sprinter` is a multi-stage method that fits the following pairwise interaction model:
$$
y = \sum_{j = 1}^p X_j \beta^\ast_j + \sum_{\ell \leq k} X_{\ell} X_k \gamma^\ast_{\ell k} + \varepsilon.
$$
This document serves as an introduction of using the package with a simple simulated data example.

## Data simulation
We consider the following simple simulation setting, where $X \sim N(\mathbf{0}, \mathbf{I}_p)$. There are two non-trivial main effects $\beta_1 = 1$, $\beta_2 = -2$, and $\beta_j = 0$ for $j > 2$. The two important interactions are $X_1 * X_3$ with $\gamma_{13} = 3$, and $X_4 * X_5$ with $\gamma_{45} = -4$. With $\varepsilon \sim N(0, 1)$, the following code simulates $n = 100$ observation from the model above with $p = 100$.
```{r}
library(sprintr)
set.seed(123)
n <- 100
p <- 100
x <- matrix(data = rnorm(n * p), nrow = n, ncol = p)
y <- x[, 1] - 2 * x[, 2] + 3 * x[, 1] * x[, 3] - 4 * x[, 4] * x[, 5] + rnorm(100)
```


# Using `sprinter` function
The function `sprinter` implements the sprinter method (please note that the function name `sprinter` is different from the package name `sprintr`), which involves the following three main steps:

- Fit a lasso (with cross-validation) of the response $y$ only on main effects $X$ (if `square = FALSE` by default) or with both main effects and squared effects $(X, X^2)$ (if `square = TRUE`).
- Carry out a screening procedure based on the residual from the previous step. The number of the selected candidate interactions can be specified by a path of `num_keep` values.
- With a path of tuning parameter `lambda`, fit a lasso of the response on main effects, squared effects (if `square = TRUE`), and selected interactions from the previous step. 

There are two tuning parameters: `num_keep` (used in Step 2) and `lambda` (used in Step 3). If `num_keep` is not specified, it will then be automatically compuated as a list of decreasing values, starting from $n / \lceil \log n \rceil$ (see, e.g., [Fan & Lv (2008)](https://orfe.princeton.edu/~jqfan/papers/06/SIS.pdf)), based on the number of tuning parameters (`n_num_keep`, default to be 5). 
If `lambda` is not specified, then `sprinter` would compute its own path of tuning parameter values based on the number of tuning parameters (`nlam`) and the range of the path (`lam_min_ratio`).
Finally, a `verbose` option (default to be `TRUE`) can be turned on to see the progress of the computation.
```{r}
fit <- sprinter(x = x, y = y, square = FALSE, nlam = 100, lam_min_ratio = 0.01)
```

## `sprinter` output
The output of `sprinter` is a `S3` object including several useful components. In particular, the component `step3` is a list of length `n_num_keep`, with `step3[[j]]` containing information from Step 3 fit when the tuning parameter in Step 2 is `num_keep[j]`.

Within each `step3` component, there exists a `idx` object representing all variables considered in Step 3:
```{r}
fit1 <- fit$step3[[1]]
fit1$idx[(p + 1) : nrow(fit1$idx), ]
```
In particular, `fit1$idx[, 1:2]` contains the indices of all the variables indices, and the last column represents their corresponding scores used to select candidate interactions in Step 2. The two columns of `idx` represents the index pair $(\ell, k)$ of a selected interaction $X_\ell * X_k$, where $\ell \leq k$.
If the first entry of an index pair is zero, i.e., $(\ell = 0, k)$, then it represents a main effect $X_k$ (with zero score).

The output `fit1$coef` is a `nrow(fit1$idx)`-by-`length(fit1$lambda)` matrix. Each column of `fit1$coef` is a vector of estimates of all variable coefficients considered in Step 3 corresponding to one value of the lasso tuning parameter `lambda`. For example, for the 30-th tuning parameter, we have the corresponding coefficient estiamte:
```{r}
estimate <- fit1$coef[, 30]
cb <- cbind(fit1$idx, estimate)
cb[cb[, 3] != 0, ]
```

## Summarizing `sprinter` output by `print`, `plot`, and `summary`
The output of `sprinter` has an associated `print` function, that prints information (the number of nonzero main effects and nonzero interactions) of Step 3 fits along a path of Step 3 tuning parameters, for a given value of Step 2 tuning parameter (specified by `which`). For example, the following codes prints the output when the 2nd value of Step-2 tuning parameter is used:
```{r}
print(fit, which = 2)
```
The `plot` function is also defined for `sprinter` output to look at the effects from a certain interaction (specified by `which`). For example, by examining the effect of interaction between $X_4$ and $X_5$, we run the following `plot` function that produces 4 panels:
```{r fig.width = 7}
plot(fit, newdata = x, index = c(1, 40), which = c(4, 5))
```
The top two panels show the marginal relationship of the predicted response on the two main effects, i.e., $\hat{\beta}_4 X_4$ and $\hat{\beta}_5 X_5$ respectively. The lower-left panel shows the dependence of the predicted response on the interaction alone, i.e., $\hat{\gamma}_{45} X_4 * X_5$, and the lower-right panel shows $\hat{\beta}_4 X_4 + \hat{\beta}_5 X_5 + \hat{\gamma}_{45} X_4 * X_5$. From the plot we see that the predicted response does not depend on either of the two main effects, but depends on their interaction (with coefficient -3.56)

Finally, `summary` function shows the dependence of coefficient estimates for each main effects (left panel) and interactions(right panel) on the Step-3 tuning parameters:
```{r fig.width = 7}
summary(fit)
```

# Using cross-validation with `cv.sprinter`
The function `cv.sprinter()` performs cross-validation to select the best value pairs of Step-2 and Step-3 tuning parameters.
```{r }
fit_cv <- cv.sprinter(x = x, y = y, square = FALSE, n_num_keep = 5, nlam = 100, lam_min_ratio = 0.01)
```

## `cv.sprinter` output
The output of `cv.sprinter` is a `S3` object. The most interesting information is `fit_cv$compact`, which is a matrix of three columns.
The first two columns show the indices pairs of all variables finally selected by cross-validation, and the last column is the coefficient estimate corresponding to those selected variables.
```{r}
fit_cv$compact
```
We see (from the first two rows and the last two rows) that the fit selected by cross-validation includes all the four important variables ($X_1, X_2, X_4*X_5, X_1 * X_3$) in the model, with relatively accurate estimates of their coefficients.

## Summarizing `cv.sprinter` output by `print` and `plot`
Associated with the output of `cv.sprinter` are the `print` and `plot` functions. `print` functions can be used to the summary of the cross-validation process, indicating information such as the best number of candidate interactions in Step 2, and the validation error mean/standard errors, number of non-zero main effects/interactions for the Step-3 tuning parameter selected by `min` rule and `1se` rule.
```{r}
print(fit_cv)
```

The `plot` function for the output of `cv.sprinter` shows the validation error across different folds as a function of Step-3 tuning parameters (for a fixed value of Step-2 tuning parameter chosen by cross-validation). The top of the plot shows the number of nonzero main effects / nonzero interactions corresponding (in orange) to a value of Step-3 tuning parameters.
```{r fig.width=7}
plot(fit_cv)
```
The blue vertical line shows the Step-3 tuning parameter selected by `min` rule, and the green vertical line shows the Step-3 tuning parameter selected by `1se` rule.

# Prediction
The `predict` function is defined for both the object returned by `sprinter` and `cv.sprinter` that computes the prediction for a new data matrix of main effects:
```{r}
newdata <- matrix(rnorm(20 * p), nrow = 20, ncol = p)
pred <- predict(fit, newdata = newdata)
```
The prediction for `sprinter` object computes the prediction at `newdata` for all the (Step-2, Step-3) tuning parameter pairs, and the prediction for `cv.sprinter` object just computes the prediction at `newdata` for the best tuning parameter pairs selected by cross-validation.
```{r}
pred_cv <- predict(fit_cv, newdata = newdata)
```

# Support for other response families
Under construction
